Decentralized state machine using Nakamoto (probabilistic) consensus\\\medskip Research Project 2022

### 2022-11-02 article/article.tex
typos

In this article, we will look at a new family of probabilitic consensus algorithms, in the context of building a decentralized state machine.

@@

The client can ask any node to add a log to the record. A log, in my implementation, is just a string, corresponding to a unique number. Both are given by the client in his request to the state machine.
If the client asks a node which is not the leader, the request is transfered to the leader. If there is already a log at the given number,
the leader returns an error message to the client. Otherwise, the leader updates its record and asks all his followers to update themselves as well.

@@

% \subsection{Byzantine fault-tolerance (BFT) with classical consensus}

% Contrary to Crash Fault, Byzantine Fault \cite{byzantine} is a type of failure where you consider that the node can fail in any possible way. In practice, it means that the node will stay up and can send malicious requests to
% interfere with network usage.

% The Raft consensus algorithm is not BFT. BFT algorithms (such as \cite{pbft}) include a layer of authentification, but are not necessarily much slower. However, there is a maximum number of malicious nodes that the algorithm can handle safely \cite{partialSynchrony}.
% For example, assuming partial synchrony of the nodes, consensus protocols can handle $t$ crash failures with $2t+1$ nodes but $t$ byzantine failures with no less than $3t+1$ nodes.

@@

The Snowball protocol is triggered when the client asks to add a new log to an index that has not been used yet. Indeed, the system works as if each index in the record corresponds to an independant Snowball run.

@@

When receiving an add-log request, the node follows this sequence, in which $logToAdd$(string) and $logPosition$(int) are variables from the request (client or other node) and $record$(map[int]string) and $cnt$(map[int]int) are variables from the node memory:
\\
\begin{algorithmic}
    \If{$record[logPosition]$ exists}
    \State reply $record[logPosition] == logToAdd$
    \Else
    \State $record[logPosition] \gets logToAdd$
    \State reply $True$
    \State initiateRequest($logToAdd$, $logPosition$)
    \EndIf
\end{algorithmic}

initiateRequest corresponds to the following sequence :

\begin{algorithmic}
    \While{$True$}
    \State $success \gets initiateQuery(req)$
    \State ($initiateQuery$ picks a sample in the network and queries it)
    \If{$success$}
    \State $cnt[logPosition]++$
    \Else
    \If{$cnt[logPosition] == 1$}
    \State delete($record$, $logPosition$)
    \State $cnt[logPosition] \gets 0$
    \Else
    \State $cnt[logPosition]--$
    \EndIf
    \EndIf
    \If{$cnt[logPosition] \geqslant CounterThreshold$}
    \State $record[logPosition] \gets logToAdd$
    \State return
    \EndIf
    \EndWhile
\end{algorithmic}

3 parameters are used in my implementation :
\begin{itemize}
    \item sampleSize (k) : the number of nodes that are queried each round
    \item majorityThreshold (α) : the number of nodes that have to reply successfully to consider that the query is successful.
          (with $\alpha > \lfloor k/2 \rfloor$)
    \item CounterThreshold (β) : the threshold of successful requests above which the node stops initiating requests and only replies with its content. (for one position)
\end{itemize}

@@

With Avalanche, to begin with, a node cannot easily change the information that is already in place among nodes. This would require the consensus to switch from one way to the other, which is nearly impossible. However, in my implementation, the client cannot either change a log at one position.
One thing the byzantine node cannot easily do in the Snowball implementation is register different informations among multiple nodes. Indeed, these nodes will want to verify the information by quering other nodes to reach consensus. The system will therefore go one way or the other.

@@

We demonstrated that it was possible to build a decentralized state machine, that could be based on a Nakamoto consensus algorithm and had interesting properties. 
In normal running state, the number of requests to reach consensus with Snowball is a function of $k * \beta$ for each node. In Raft, the leader must handle $n-1$ request on its own. 
Raft has one consensus for the whole system, whereas in Snowball to each value corresponds a different consensus run. For most applications, classical consensus is probably the way to go. Still, it would be useful to lead some more research on the topic. 
In particular, it would be interesting to approximate the value of ε $\forall n, k, \beta$. 
A comparison focused on Byzantine Fault Tolerance would also be interesting. To do that one would need to use PBFT \cite{pbft} instead of the Raft.

>>>>

In this article, we will look at a new family of probabilistic consensus algorithms, in the context of building a decentralized state machine.

@@

The client can ask any node to add a log to the record. A log, in my implementation, is just a string, corresponding to a unique number. Both are given by the client in his request to the state machine.
If the client asks a node which is not the leader, the request is transferred to the leader. If there is already a log at the given number,
the leader returns an error message to the client. Otherwise, the leader updates its record and asks all his followers to update themselves as well.

@@

@@

The Snowball protocol is triggered when the client asks to add a new log to an index that has not been used yet. Indeed, the system works as if each index in the record corresponds to an independent Snowball run.

@@

When receiving an add-log request, the node follows this sequence, in which $logToAdd$(string) and $logPosition$(int) are variables from the request (client or other node) and $record$(map[int]string) and $cnt$(map[int]int) are variables from the node memory:
\\
\begin{algorithmic}
    \If{$record[logPosition]$ exists}
    \State reply $record[logPosition] == logToAdd$
    \Else
    \State $record[logPosition] \gets logToAdd$
    \State reply $True$
    \State initiateRequest($logToAdd$, $logPosition$)
    \EndIf
    \\
\end{algorithmic}
initiateRequest corresponds to the following sequence :
\\
\begin{algorithmic}
    \While{$True$}
    \State $success \gets initiateQuery(req)$
    \State ($initiateQuery$ picks a sample in the network and queries it)
    \If{$success$}
    \State $cnt[logPosition]++$
    \Else
    \If{$cnt[logPosition] == 1$}
    \State delete($record$, $logPosition$)
    \State $cnt[logPosition] \gets 0$
    \Else
    \State $cnt[logPosition]--$
    \EndIf
    \EndIf
    \If{$cnt[logPosition] \geqslant CounterThreshold$}
    \State $record[logPosition] \gets logToAdd$
    \State return
    \EndIf
    \EndWhile
    \\
\end{algorithmic}
3 parameters are used in my implementation :
\begin{itemize}
    \item sampleSize (k) : the number of nodes that are queried each round
    \item majorityThreshold (α) : the number of nodes that have to reply successfully to consider that the query is successful.
          (with $\alpha > \lfloor k/2 \rfloor$)
    \item CounterThreshold (β) : the threshold of successful requests above which the node stops initiating requests and only replies with its content. (for one position)
\end{itemize}

@@

With Avalanche, to begin with, a node cannot easily change the information that is already in place among nodes. This would require the consensus to switch from one way to the other, which is nearly impossible. However, in my implementation, the client cannot either change a log at one position.
One thing the byzantine node cannot easily do in the Snowball implementation is register different data among multiple nodes. Indeed, these nodes will want to verify the information by querying other nodes to reach consensus. The system will therefore go one way or the other.

@@

We demonstrated that it was possible to build a decentralized state machine, that could be based on a Nakamoto consensus algorithm and had interesting properties.
In normal running state, the number of requests to reach consensus with Snowball is a function of $k * \beta$ for each node. In Raft, the leader must handle $n-1$ request on its own.
Raft has one consensus for the whole system, whereas in Snowball to each value corresponds a different consensus run. For most applications, classical consensus is probably the way to go. Still, it would be useful to lead some more research on the topic.
In particular, it would be interesting to approximate the value of ε $\forall n, k, \beta$.
A comparison focused on Byzantine Fault Tolerance would also be interesting. To do that, one would need to use PBFT \cite{pbft} instead of the Raft.
